---
title: "Lab 06 - Functions and iterations"
author: "Jessica Preston"
date: "`r Sys.Date()`"
output: html_document
---


### Packages and Data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(ggplot2)
```

```{r diamon_rename}
diamonds <- diamonds %>% rename(
    depth_pct = depth,
    length_mm = x,
    width_mm = y,
    depth_mm = z,
    table_pct = table
  )
```


## Exercise 1

```{r}
# Do some preliminary investigations & cleaning

ggplot(data = diamonds,
       mapping = aes(x = width_mm,
                      y = depth_mm)) +
  geom_point()

diamonds <- diamonds %>%
  filter(length_mm != 0, length_mm < 20, width_mm != 0, width_mm < 20, depth_mm != 0, depth_mm < 20)

```


## Exercise 2

```{r}

diamonds <- diamonds %>%
  mutate(table_mm = table_pct * width_mm / 100)

# Overall table_mm arithmetic mean?

diamonds %>%
  summarise(mean_table_length = mean(table_mm, na.rm = TRUE))




# table_mm arithmetic mean within each clarity category

diamonds %>%
  group_by(clarity) %>%
  summarise(average_table_length = mean(table_mm, na.rm = TRUE))




```

_As the clarity increases, the average table length decreases._

## Exercise 3

_Do derivation on paper, unless you know how to write LaTeX code_


## Exercise 4

```{r}
# Write gmean() command


gmean <- function(x){
  n = length(x)
  exp(1/n * mean(log(x)))
}

gmean(diamonds$table_mm)

```

## Exercise 5

```{r, error = TRUE}

test_data <- list(
  test1 = c(2.1, 3.8, 4.2),
  test2 = c(1, 10, 100, 1000),
  test3 = c(0, 1, 4),
  test4 = c(0.38,  0.94, -1.56),
  test5 = c(TRUE, TRUE),
  test6 = c("6", "7", "8")
)


# Create for loop to test gmean() on the above examples

for (i in seq_along(test_data)){
  print(gmean(test_data[[i]]))
}

```


## Exercise 6

```{r}
# Copy gmean() from Ex4 and edit it to check the input

gmean <- function(x){
  
  if(!is.numeric(x)){
    warning("Use numeric values only")
    return(NaN)
  }
  
  if( any(x <= 0) ){
    warning("Use positive numbers only")
    return(NaN)
  }
  
  
  
  n = length(x)
  exp(1/n * mean(log(x)))
}



# Copy your code from Ex 5 to test your new command

for (i in seq_along(test_data)){
  print(gmean(test_data[[i]]))
}



```


## Exercise 7

```{r}
# Overall table_mm geometic mean?

gmean(diamonds$table_mm[[i]])




# table_mm geometic mean within each clarity category

diamonds %>%
  group_by(clarity) %>%
  summarise(
    table_mean = mean(table_mm),
    table_median = median(table_mm),
    table_gmean = gmean(table_mm[[i]])
  )




```

_The table shows that as clarity increases, the average size of the table decrease for arithmetic mean length and median length, whilst the clarity and geometric mean have a non-linear relationship._


## Exercise 8 (optional)

```{r}
# Create gmean2() that computes the geometric mean 
#   that uses the original definition






```


## Exercise 9 (optional)

```{r}
# Create a function that computes the harmonic mean







```
